---
title: "World War II Allied Aerial Bombing Operations (THOR Database)"
author: "Mohamed Dhia Hammami"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(DBI)
library(RSQLite)
library(tidyverse)
library(leaflet)
```

## 1. Database Connection

We connect directly to the pre-indexed SQLite database containing the complete 178,281 THOR attack records.

```{r}
con <- dbConnect(RSQLite::SQLite(), "../data/processed/thor_wwii.sqlite")
dbListTables(con)
```

## 2. Total Bomb Tonnage by Theater of War

```{r}
theaters <- dbGetQuery(con, "
  SELECT 
    THEATER as theater, 
    COUNT(*) as missions,
    ROUND(SUM(total_tons_clean), 1) as total_tons,
    ROUND(SUM(CASE WHEN tonnage_outlier_reason IS NULL THEN TONS_OF_HE END), 1) as he_tons,
    ROUND(SUM(CASE WHEN tonnage_outlier_reason IS NULL THEN TONS_OF_IC END), 1) as ic_tons
  FROM missions
  GROUP BY THEATER
  ORDER BY total_tons DESC
")

knitr::kable(theaters, caption = "WWII Aerial Bombing Tonnage by Theater")
```

```{r fig.width=9, fig.height=5}
ggplot(theaters, aes(x = reorder(theater, total_tons), y = total_tons, fill = theater)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Total Bomb Tonnage by Theater of Operations",
       x = "Theater", y = "Total Tons of Bombs") +
  theme_minimal() +
  theme(legend.position = "none")
```

## 3. Tunisia & North Africa Campaign (1942–1943)

Focusing on the Battle of Tunisia and Mediterranean interdiction strikes.

```{r}
tunisia_targets <- dbGetQuery(con, "
  SELECT 
    target_location,
    avg_lat as lat,
    avg_lon as lon,
    mission_count,
    total_tons,
    first_mission,
    last_mission
  FROM targets_summary
  WHERE UPPER(target_country) = 'TUNISIA'
  ORDER BY total_tons DESC
")

knitr::kable(head(tunisia_targets, 15), caption = "Top 15 Bombing Targets in Tunisia")
```

### Interactive Map of Tunisia Strikes

```{r fig.width=10, fig.height=7}
leaflet(tunisia_targets) %>%
  addProviderTiles("CartoDB.DarkMatter") %>%
  setView(lng = 10.0, lat = 35.5, zoom = 7) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = ~pmax(4, pmin(25, sqrt(total_tons) * 0.4)),
    color = "#f97316",
    fillOpacity = 0.7,
    popup = ~paste0("<b>", target_location, "</b><br>",
                   "Missions: ", format(mission_count, big.mark = ","), "<br>",
                   "Total Tons: ", format(round(total_tons, 1), big.mark = ","), " t<br>",
                   "First Strike: ", first_mission, "<br>",
                   "Last Strike: ", last_mission)
  )
```

## 4. Top Bombing Targets in Europe (ETO)

```{r}
eto_targets <- dbGetQuery(con, "
  SELECT target_country, target_location, mission_count, total_tons, first_mission, last_mission
  FROM targets_summary
  WHERE theater = 'ETO'
  ORDER BY total_tons DESC
  LIMIT 15
")

knitr::kable(eto_targets, caption = "Top 15 Strategic Bombing Targets in Europe")
```

## 5. Aircraft Effectiveness & Mission Utilization

```{r}
aircraft_stats <- dbGetQuery(con, "
  SELECT 
    aircraft_full_name,
    aircraft_category,
    COUNT(*) as missions,
    ROUND(SUM(total_tons_clean), 1) as total_tons,
    ROUND(AVG(total_tons_clean), 2) as avg_tons_per_mission
  FROM missions
  WHERE aircraft_full_name IS NOT NULL
  GROUP BY aircraft_full_name, aircraft_category
  HAVING missions >= 1000
  ORDER BY total_tons DESC
  LIMIT 10
")

knitr::kable(aircraft_stats, caption = "Top 10 Aircraft by Bomb Tonnage Delivered")
```

```{r}
dbDisconnect(con)
```
